### 2017-02-04 intro.tex
fresh changes

%!TEX root =  autocontgrlp.tex
\section{Introduction}
Markov decision processes (MDPs) are a powerful mathematical framework to study optimal sequential decision making problems  arising in science and engineering. In an MDP, the configuration of the system is described by the state variables (or simply the state) and the system evolves in a stochastic manner which depends on the action made. The set of all states is denoted by $S$, the state space and the set of all actions is denoted by $A$, the action space.
%An instance of an MDP is a controlled queue setting, where there is a cost associated with the number of customers in the queue, and the aim is to control the service level depending on the number of %customers so as to achieve a minimum cumulative cost. In more general terms, given any MDP,
We are interested in computing what is known as the optimal policy $u^*$ (a map from $S$ to $A$). The so-called dynamic programming (DP) methods first compute what is known as the optimal \emph{value-function} ($J^*$), a vector whose dimension is the number of states, and use it to compute $u^*$. When the number of state is small conventional DP techniques, such as value-, or policy-iteration, or linear programming (LP) can be used to compute $J^*$ and $u^*$\cite{BertB}.\par
As the number of states is grows, it becomes increasingly difficult to compute the exact values of $J^*$ and $u^*$. A practical solution then is to compute an approximate value function $\tilde{J}$ instead of $J^*$. Approximate dynamic programming (ADP) methods combine an approximation architecture to represent $\tj$ and a conventional DP method to compute $\tj$. Eventually, ADP methods output a sub-optimal policy $\tu$ using the $\tj$ they compute. %Here, success depends on the quality of approximation, i.e., on the quantity $||J^*-\tilde{J}||$ for an appropriately chosen norm.\par
Linear function approximation (LFA), i.e., letting $\tilde{J}=\Phi r^*$ where $\Phi$ is a so-called feature matrix and $r^*$ is a weight vector to be computed, is the most widely used method of approximation. Here, dimensionality reduction is achieved by choosing $\Phi$ to have fewer columns in comparison to the number of states, holding the promise of being able to work with MDPs regardless of the number of states.\par
It is natural to expect that approximations lead to errors and it is important to quantify the errors. For a given ADP method, theoretical performance analysis analytically bounding the error terms $||J^*-\tilde{J}||$  and $\norm{J^*-J_{\tu}}$ which denote the error in approximating the value function, and performance loss due to following policy $\tu$ (here $J_{\tu}$ is the value of $\tu$) respectively. Further, in most cases the error terms reveal some structure that can offer insights and act as guide to the designer of the ADP method (for example the choice of $\Phi$). \par

%While many ADP methods use LFA, not all of them are successful. For instance, ADP methods that use linear least squares projection are known to exhibit `policy oscillations' \cite{dpchapter}, i.e., %output a repeating sequence of bad sub-optimal policies.
%Such an analysis is important to establish that the error is always bounded.
%Further, in most cases the error terms reveal some structure that can offer insights and act as guide to the designer of the ADP method (for example the choice of $\Phi$). \par

The \emph{approximate linear program} (ALP) \cite{ALP,CS,SALP,ALP-Bor} and its variants combine LFA and the LP approach. Theoretical performance analysis of the ALP can be found in \cite{ALP}.
%and a salient feature is that it does not suffer from issues such as `policy oscillations'\footnote{ADP methods that use linear least squares projection are known to exhibit `policy oscillations' %\cite{BertB}, i.e., output a repeating sequence of bad sub-optimal policies. Since our focus in this paper is the ALP formulation, we refrain from a detailed presentation of the other methods.}

A critical shortcoming of vanilla ALP is that the number of constraints are of the order of the size of the state space, making this vanilla version intractable. Two approaches have been found empircally successful in addressing the issue of large number of constraints in the ALP. In the first approach, a random subset of constraints is chosen (dropping the rest), thereby formulating a \emph{reduced linear program} (RLP). The performance analysis of the RLP can be found in \cite{CS}, however, the bounds hold only in high probability under idealized assumptions. The second approach involves employing function approximation in the dual variables of the ALP \cite{ALP-Bor,dolgov}. However, to this date, there exist no theoretical guarantees bounding the loss in performance due to this approach.\par
Our motivation stems from the fact that ALP with tractable number of constraints will result in a full dimeniosnality free ADP method. However, constraint reduction in the ALP is an extra source of error (in addition to the error due to LFA), which has not been theoretically well understood.  The focus of this paper is to fill this gap in theory by deriving performance bounds for constraint reduction in the ALP formulation.
The salient aspects of our contribution are listed below:
\begin{enumerate}
\item We define a generalized reduced linear program (GRLP) which has a tractable number of constraints that are obtained as positive linear combinations of the original constraints. The GRLP amounts to linear function approximation of the dual variables, and the RLP is a special case of GRLP.
		\item We develop novel analytical machinery to relate $\hat{J}$, the solution to the GRLP, and the optimal value function $J^*$ by bounding the prediction error $||J^*-\hj||$ (\Cref{cmt2mn}). 
		\item We also bound the performance loss due to using the policy $\hu$ that is one-step greedy with respect to $\hj$ (Theorem~\ref{polthe}).
		\item Our analysis is based on two novel $\max$-norm contraction operators and our results hold \emph{deterministically}, as opposed to the results on RLP \cite{SALP,CS}, where the guarantees have a probabilistic nature.
%Our analysis also makes use of arguments based on \emph{Lyanpunov} function, an approach much similar to prior works in ALP literature \cite{ALP,SALP}.
\item Our results on the GRLP are the first to theoretically analyze the use of linear function approximation of Langrangian (dual) variables underlying the constraints.
\item A numerical example in controlled queues is provided to illustrate the theory.
\end{enumerate}
%ALP with tractable number of constraints would result in  a full dimensionality free ADP method, without issues such as policy oscillations.
A short and preliminary version of this paper without the theoretical analysis is available in \cite{aaaipaper}.

>>>>

%!TEX root =  autocontgrlp.tex
\section{Introduction}
The  framework of Markov decision processes (MDPs) is useful to mathematically cast optimal sequential decision making problems arising in science and engineering. By casting a given decision problem as an MDP, the goal then is to find an optimal policy ($u^*$) which is the best decision rule. The so-called dynamic programming (DP) methods find an optimal policy by computing the optimal \emph{value-function} ($J^*$), a vector whose dimension is the number of states. The value function $J^*$ stands for the cumulative return of making decisions according to $u^*$. Conventional DP  techniques, such as value-, or policy-iteration, or linear programming (LP) \cite{BertB} can compute the exact value of $J^*$. However, a shortcoming of these conventional methods is that their computational overhead grows with the number of states, a practical hindrance when the MDP has a large number of states.\par
In this paper we consider the problem of using dynamic programming with MDPs with large state spaces. A practical way to tackle the issue of large number of states is to compute an approximate value function $\tilde{J}$ instead of $J^*$, and a sub-optimal policy $\tu$ can be obtained using $\tj$. The sub-optimal policy can then be used to make the decision and results in a cumulative return given by $J_{\tu}$. The idea here is forsake optimality to gain computational tractability. The success depends on the quality of approximation, i.e., on the quantities $||J^*-\tilde{J}||$ and $\norm{J^*-J_{\tu}}$ ($\norm{\cdot}$ is an appropriate norm) known as prediction error and the control error respectively.\par
Linear function approximation (LFA), i.e., letting $\tilde{J}=\Phi r^*$ where $\Phi$ is a so-called feature matrix and $r^*$ is a weight vector to be computed, is the most widely used method of approximation. Here, dimensionality reduction is achieved by choosing $\Phi$ to have fewer columns in comparison to the number of states, holding the promise of being able to work with MDPs regardless of the number of states.\par
The \emph{approximate linear program} (ALP) \cite{ALP,CS,SALP,ALP-Bor} and its variants introduce linear function approximation in the linear programming formulation. A critical shortcoming of vanilla ALP is that the number of constraints are of the order of the size of the state space, making this vanilla version intractable. A way out is to choose a subset of constraints at random and drop the rest, thereby formulating a \emph{reduced linear program} (RLP). The performance analysis of the RLP can be found in \cite{CS} and the RLP has also been shown to perform well in experiments \cite{ALP,CS,CST}. An alternative approach to handle the issue of large number of constraints is to employ function approximation in the dual variables of the ALP \cite{ALP-Bor,dolgov}, an approach that was also found useful in experiments. However, to this date, there exist no theoretical guarantees bounding the loss in performance resulting from such an approximation.
In this paper, we generalize the RLP to define a generalized reduced linear program (GRLP) which has a tractable number of constraints that are obtained as positive linear combinations of the original constraints.
The salient aspects of our contribution are listed below:
\begin{enumerate}
\item We develop novel analytical machinery to relate $\hat{J}$, the solution to the GRLP, and the optimal value function $J^*$ by bounding the prediction error $||J^*-\hj||$ (\Cref{cmt2mn}).
\item We also bound the performance loss due to using the policy $\hu$ that is one-step greedy with respect to $\hj$ (Theorem~\ref{polthe}).
\item Our analysis is based on two novel $\max$-norm contraction operators and our results hold \emph{deterministically}, as opposed to the results on RLP \cite{SALP,CS}, where the guarantees have a probabilistic nature.
\item Our analysis also makes use of arguments based on \emph{Lyapunov} function, an approach much similar to prior works in ALP literature \cite{ALP,SALP}.
\item Our results on the GRLP are the first to theoretically analyze the use of linear function approximation of Lagrangian (dual) variables underlying the constraints.
\item A numerical example in controlled queues is provided to illustrate the theory.
\end{enumerate}
A short and preliminary version of this paper without the theoretical analysis is available in \cite{aaaipaper}.
\begin{comment}
\section{Introduction}
Markov decision processes (MDPs) are a powerful mathematical framework to study optimal sequential decision making problems  arising in science and engineering. In an MDP, the configuration of the system is described by the state variables (or simply the state) and the system evolves in a stochastic manner which depends on the action made. The set of all states is denoted by $S$, the state space and the set of all actions is denoted by $A$, the action space.
%An instance of an MDP is a controlled queue setting, where there is a cost associated with the number of customers in the queue, and the aim is to control the service level depending on the number of %customers so as to achieve a minimum cumulative cost. In more general terms, given any MDP,
An optimal policy $u^*$ is a map from $S$ to $A$ that results in the best cumulative reward that can be obtained by starting from any state. The rewards for the various The so-called dynamic programming (DP) methods first compute what is known as the optimal \emph{value-function} ($J^*$), a vector whose dimension is the number of states, and use it to compute $u^*$. When the number of state is small conventional DP techniques, such as value-, or policy-iteration, or linear programming (LP) can be used to compute $J^*$ and $u^*$\cite{BertB}.\par
Curse-of-dimensionality  refers to the fact that the number of states grows exponentially in the number of state variables. Many practical systems such as controlled queues, inventory control etc are
afflicted by the curse, i.e., have large number of states. In such scenarios, it becomes increasingly difficult to compute the exact values of $J^*$ and $u^*$ because the DP methods are based on computations involving as many (or even more) number of variables as the number of states. A practical solution then is to compute an approximate value function $\tilde{J}$ instead of $J^*$. Approximate dynamic programming (ADP) methods combine an approximation architecture to represent $\tj$ and a conventional DP method to compute $\tj$. Eventually, ADP methods output a sub-optimal policy $\tu$ using the $\tj$ they compute. %Here, success depends on the quality of approximation, i.e., on the quantity $||J^*-\tilde{J}||$ for an appropriately chosen norm.\par
Linear function approximation (LFA), i.e., letting $\tilde{J}=\Phi \tr$ where $\Phi \in \R^{|S|\times k}$ is a so-called feature matrix and $\tr*\in \R^k$ is a weight vector (to be computed), is the most widely used method of approximation. Here, dimensionality reduction is achieved by choosing $\Phi$ to have fewer columns in comparison to the number of states ($k<<|S|$), holding the promise of being able to work with MDPs regardless of the number of states.\par
It is natural to expect that approximations lead to errors and it is important to quantify the errors. For a given ADP method, theoretical performance analysis analytically bounding the error terms $||J^*-\tilde{J}||$  and $\norm{J^*-J_{\tu}}$ which denote the error in approximating the value function, and performance loss due to following policy $\tu$ (here $J_{\tu}$ is the value of $\tu$) respectively. Further, in most cases the error terms reveal some structure that can offer insights and act as guide to the designer of the ADP method (for example the choice of $\Phi$). \par
%While many ADP methods use LFA, not all of them are successful. For instance, ADP methods that use linear least squares projection are known to exhibit `policy oscillations' \cite{dpchapter}, i.e., %output a repeating sequence of bad sub-optimal policies.
%Such an analysis is important to establish that the error is always bounded.
%Further, in most cases the error terms reveal some structure that can offer insights and act as guide to the designer of the ADP method (for example the choice of $\Phi$). \par

The \emph{approximate linear program} (ALP) \cite{ALP,CS,SALP,ALP-Bor} and its variants introduce LFA in the linear programming formulation to dynamic programming. Theoretical performance analysis of the ALP can be found in \cite{ALP}.
%and a salient feature is that it does not suffer from issues such as `policy oscillations'\footnote{ADP methods that use linear least squares projection are known to exhibit `policy oscillations' %\cite{BertB}, i.e., output a repeating sequence of bad sub-optimal policies. Since our focus in this paper is the ALP formulation, we refrain from a detailed presentation of the other methods.}
The number of variables of the ALP is only $k$, a critical shortcoming of the ALP is that the number of constraints are of the order of the size of the state space, making it intractable.
Two approaches have been found empirically successful \cite{CS,dolgov,ALP-Bor} in addressing the issue of large number of constraints in the ALP. In the first approach \cite{CS}, a random subset of constraints is chosen (dropping the rest), thereby formulating a \emph{reduced linear programming} (RLP) problem. The performance analysis of the RLP can be found in \cite{CS}, however, the bounds hold only in high probability under idealized assumptions. The second approach involves employing function approximation in the dual variables of the ALP \cite{ALP-Bor,dolgov}. However, to this date, there exist no theoretical guarantees bounding the loss in performance due to this approach.\par

Our motivation stems from the fact that ALP with tractable number of constraints will result in a full dimeniosnality free ADP method. However, constraint reduction in the ALP is an extra source of error (in addition to the error due to LFA), which has not been theoretically well understood.  The focus of this paper is to fill this gap in theory by deriving performance bounds for constraint reduction in the ALP formulation.
The salient aspects of our contribution are listed below:
\begin{enumerate}
\item We define a generalized reduced linear programming (GRLP) which has a tractable number of constraints that are obtained as positive linear combinations of the original constraints. The GRLP amounts to linear function approximation of the dual variables, and the RLP is a special case of GRLP.
\item We develop a novel analytical machinery to bound the prediction error $\norm{J^*-\hj}$ where $hj$ is the solution to the GRLP.
		\item We also bound the performance loss due to using the policy $\hu$ that is one-step greedy with respect to $\hj$ (Theorem~\ref{polthe}).
		\item Our analysis is based on two novel $\max$-norm contraction operators and our results hold \emph{deterministically}, as opposed to the results on RLP \cite{SALP,CS}, where the guarantees have a probabilistic nature.
%Our analysis also makes use of arguments based on \emph{Lyapunov} functions, an approach much similar to prior works in the ALP literature \cite{ALP,SALP}.
\item Our results on GRLP are the first to theoretically analyze the use of linear function approximation of Lagrangian (dual) variables underlying the constraints.
\item A numerical example in controlled queues is provided to illustrate the theory.
\end{enumerate}
%ALP with tractable number of constraints would result in  a full dimensionality free ADP method, without issues such as policy oscillations.
A short and preliminary version of this paper without the theoretical analysis is available in \cite{aaaipaper}. The rest of the paper is organized as follows:
\end{comment}

### 2017-02-04 mdp.tex
fresh changes

\section{Markov Decision Processes (MDPs) and Linear Programming}
We consider MDPs with a finite state space $S=\{1,2,\ldots,n\}$ and finite action space $A=\{1,2,\ldots,d\}$, where the sentiment is that $n$ is large. For simplicity, we assume that all actions are feasible in all states. The probability transition kernel $P$ collects the probabilities $p_a(s,s')$ of transitioning from state $s$ to state $s'$ under the action $a$ for all possible $s,s'\in S$ and $a\in A$. We denote the reward (or gain) obtained for performing action $a\in A$ in state $s\in S$ by $g_a(s)$.
A stationary deterministic policy\footnote{For the scope of this paper, it suffices to restrict our attention to stationary deterministic policies.} (SDP), or simply a policy, is a map $u\colon S\ra A$ that specifies for each state what action to select in that state.
Given an SDP $u$, the expected total discounted reward corresponding to starting at state $s$, while choosing actions as dictated by $u$ for the states encountered, is

@@

J_u(s)\stackrel{\Delta}{=}\E[\sum_{n=0}^\infty \alpha^n g_{a_n}(s_n)|s_0=s,a_n=u(s_n)\mbox{ }\forall n\geq 0],\nn

@@

where $\alpha \in (0,1)$ is the so-called discount factor and $s_{n+1} \sim p_{a_n}(s_n,\cdot)$, $n\ge 0$. 

@@

is called the value function underlying policy $u$.
The \emph{optimal policy} $u^*$ is one that in each state $s\in S$ achieves the best possible total expected discounted reward from that state. That is, $J^{u^*}(s) = J^*(s) \eqdef \max_{u\in U} J_u(s)$
where $U$ is the set of all SDPs and $J^*$ is coined the \emph{optimal value function}.%
\footnote{In our case an optimal (SDP) $u^*$ exists and is well defined \cite{BertB}.}, 
Any optimal policy $u^*$ and value function $J^*$ obey the Bellman equation (BE): for all $ s \in S$,

@@

\label{bellval}J^*(s)&=\max_{ a\in A}\big(g_a(s)+\alpha \sum_{s'}p_a(s,s')J^*(s')\big),\\
\label{bellpol}u^*(s)&= \underset{a\in A}{\argmax}\big(g_a(s)+\alpha \sum_{s'}p_a(s,s')J^*(s')\big)\,.

@@

Once $J^*$ is computed, $u^*$ can be obtained via \eqref{bellpol}
for each state relatively cheaply (e.g., having the ability to sample from the next-state distribution at
a given state-action pair).
Thus computing the value function is at the heart of most solution methods to MDP.\par
%\subsection{Linear Programming}
As it is well known, the optimal value function $J^*$ can be obtained by solving the following linear program:
\begin{align}\label{mdplp}
\begin{split}
\min_{J\in \R^n}\, &c^\top J\,\,\text{s.t.}\\
%\mb
J(s)&\geq g_a(s)+\alpha\sum_{s'}p_a(s,s')J(s'),\,\,
\forall (s,a)\in S\times A,
\end{split}
where $c$ is \emph{any} vector of $\R^n$ such that $c(s)>0, \forall s \in S$. Without loss of generality, we may and will assume in what follows that $\sum_{s\in S}c(s)=1$, i.e., $c$ is a probability distribution over $S$.\par
%\subsection{Notation}
The value functions $J^u$ or $J^*$ are elements of $\R^S$.
In what follows it will be useful for us to treat these as $n$-dimensional vectors, i.e., elements of $\R^n$, effectively identifying $\R^S$ with $\R^S$ in the natural way. Similarly we identify $\R^{nd}$ with $\R^{S\times A}$.
The following definitions will be useful later:
%The Bellman operator $T$ is defined using the model parameters of the MDP as follows:
\begin{definition}Define

@@

\item The Bellman operator $T\colon \R^n \ra \R^n$ as $(TJ)(s)=\max_{a \in A}\big(g_a(s)+\alpha \sum_{s'} p_a(s,s')J(s')\big).

@@

\item \label{bellactval} The Bellman operator (of action values) $H: \R^n \to \R^{nd}$ for state-action values as $HJ=[ H_1 J,\cdots,H_d J]^\top\in \R^{nd},$ where $(H_a J)(s)= g_a(s)+\alpha \sum_{s'}p_a(s,s') J(s')$.
\item
||J||_{\infty,\rho}=\max_{s \in S} \frac{|J(s)|}{\rho(s)}$.
\item The discounted maximal inflation of $\chi$ due to $P = (p_a)_{a\in A}$ as $\beta_{\chi}=\max_{s \in S} \frac{\underset{a \in A}{\max}\big(\alpha\sum_{s'}p_a(s,s')\chi(s')\big)}{\chi(s)}$.
\item The function $\chi:S\to\R_+$ above to be a \emph{Lyapunov} function for $P = (p_a)_{a\in A}$ if $\beta_{\chi}<1$.
\item $E$ to be the $nd\times n$ matrix given by $E=[I,\ldots,I]^\top$, i.e., $E$ is obtained by stacking $d$ identical $n\times n$ identity matrices one over the other.
\item A policy $u_J$ is said to be greedy with respect to (w.r.t.) $J\in \R^n$ if for any $s\in S$,

@@

\begin{align*}u_J(s)\eqdef\underset{a\in A}{\argmax}\big(g_a(s)+\alpha \sum_{s'}p_a(s,s')J(s')\big).\end{align*}
\item The weighted $L_1$-norms $\norm{\cdot}_{1,c}$ with respect to a probability distribution $c$ as $

@@

\item $J_1\leq J_2$, when $J_1(s)\leq J_2(s),~\forall s\in S$.
\item $\one\in \R^n$ to be the vector whose all coordinates are $1$.

@@

>>>>

\section{Background: Markov Decision Processes (MDPs)}
The framework of Markov decision processes (MDPs) is useful to mathematically cast optimal sequential decision making problems in stochastic environments. In this section, we present a brief overview of the MDP framework, introducing the notions of policy, value function and optimality (please refer to \cite{BertB} for a detailed presentation on MDP).\par
\textbf{Model:} We consider an MDP with a finite state space given by $S=\{1,2,\ldots,n\}$ and a finite action space given by $A=\{1,2,\ldots,d\}$ (where the sentiment is that $n$ is large). For simplicity, we assume that all actions are feasible in all states. The probability transition kernel $P$ collects the probabilities $p_a(s,s')$ of transitioning from state $s$ to state $s'$ under the action $a$ for all possible $s,s'\in S$ and $a\in A$. We denote the reward (or gain) obtained for performing action $a\in A$ in state $s\in S$ by $g_a(s)$.\par
\textbf{Policy:} A stationary deterministic policy\footnote{For the scope of this paper, it suffices to restrict our attention to stationary deterministic policies.} (SDP), or simply a policy, is a map $u\colon S\ra A$ that specifies for each state what action to select in that state. Under an SDP, an MDP is a Markov chain whose probability transition matrix is denoted by $P_u$.\par
\textbf{Value Function:} Given an SDP $u$, the expected total discounted reward corresponding to starting at a state $s\in S$ at $t=0$, while choosing actions as dictated by $u$ for the states encountered in the future (i.e., $t>0$), is

@@

J_u(s)\stackrel{\Delta}{=}\E[\sum_{t=0}^\infty \alpha^t g_{a_t}(s_t)|s_0=s,a_t=u(s_t)\mbox{ }\forall t\geq 0],\nn

@@

where $\alpha \in (0,1)$ is the so-called discount factor and $s_{t+1} \sim p_{a_t}(s_t,\cdot)$, $t\ge 0$.

@@

is called the value function underlying policy $u$. The value functions $J_u$ is an elements of $\R^S$. In what follows it will be useful for us to treat it as an $n$-dimensional vector, i.e., an element of $\R^n$, effectively identifying $\R^S$ with $\R^n$ in the natural way. Similarly we identify $\R^{nd}$ with $\R^{S\times A}$, where $d=|A|$.\par
\textbf{Optimality and the Bellman Equation:} The \emph{optimal policy} $u^*$ is one that in each state $s\in S$ achieves the best possible total expected discounted reward from that state. That is, $J_{u^*}(s) = J^*(s) \eqdef \us{\max}{u\in U} J_u(s)$
where $U$ is the set of all SDPs and $J^*$ is coined the \emph{optimal value function}.
\footnote{In our case an optimal (SDP) $u^*$ exists and is well defined \cite{BertB}.}
Any optimal policy $u^*$ and value function $J^*$ obey the Bellman equation (BE) which states that for all $ s \in S$,

@@

\label{bellval}J^*(s)&=\max_{ a\in A}\big(g_a(s)+\alpha \us{\sum}{s'\in S}p_a(s,s')J^*(s')\big),~\text{and}\\
\label{bellpol}u^*(s)&= \underset{a\in A}{\argmax}\big(g_a(s)+\alpha \us{\sum}{s'\in S}p_a(s,s')J^*(s')\big),\,.

@@

where ties in \eqref{bellpol} are resolved arbitrarily.\par
We now present some definition which will be useful later:
\begin{definition}\label{notations}

@@

\item The Bellman operator $T\colon \R^n \ra \R^n$ is given by $(TJ)(s)=\max_{a \in A}\big(g_a(s)+\alpha \sum_{s'} p_a(s,s')J(s')\big).

@@

\item \label{bellactval} The Bellman operator (of action values) $H: \R^n \to \R^{nd}$ for state-action values is given by $HJ=[ H_1 J,\cdots,H_d J]^\top\in \R^{nd},$ where $(H_a J)(s)= g_a(s)+\alpha \sum_{s'}p_a(s,s') J(s')$.
\item\label{emat} The $nd\times n$ matrix $E$ is given by $E=[I,\ldots,I]^\top$, i.e., $E$ is obtained by stacking $d$ identical $n\times n$ identity matrices one over the other.
\item\label{greedy} A policy $u_J$ is said to be greedy with respect to (w.r.t.) $J\in \R^n$ if for any $s\in S$,

@@

\begin{align*} u_J(s)\eqdef\underset{a\in A}{\argmax}\big(g_a(s)+\alpha \us{\sum}{s'\in S} p_a(s,s')J(s')\big).\end{align*}
\item\label{norms} The weighted $L_1$-norms $\norm{\cdot}_{1,c}$ with respect to a probability distribution $c$ is given by $

@@

\item For $J_1, J_2\in \R^n$ we write $J_1\leq J_2$ when $J_1(s)\leq J_2(s),~\forall s\in S$.
\item We define $\one\in \R^n$ to be the vector whose coordinates are all equal to $1$.

@@

We would like to point out that from \Cref{notations}, it follows that for any $J\in \R^n$ the condition $J\geq TJ$ can be rephrased as $EJ\geq HJ$.
\subsection{Dynamic Programming: Objective and Methods}
Once a given sequential decision making is cast in the MDP framework, it remains for us to still compute the optimal policy $u^*$. The idea behind Dynamic Programming (DP) methods is to use the BE \eqref{bellval} to first compute $J^*$, and then a near-optimal action $u^*$ can be obtained via \eqref{bellpol}
for each state relatively cheaply even for large state-spaces (e.g., having the ability to sample from the next-state distribution at
a given state-action pair). Thus computing $J^*$ is at the heart DP methods \cite{BertB} such as value-, or policy-iteration and linear programming (LP) \eqref{mdplp} which can be used in practice when the MDP has a tractable number of states. However, when the MDP has a large number of states, using DP methods to obtain $J^*$ is tedious, since they involve computations in variables that are of the order of the number of states.\par
Approximate dynamic programming (ADP) methods are based on DP methods, but employ (value) function approximation to ease the computational overhead encountered in MDPs with a large number of states. In particular, linear function approximation (LFA) has been used widely, wherein, $J^*$ is approximated by $J^*\approx\tj =\Phi \tr$, where $\Phi$ is an $n\times k$ feature matrix and $\tr\in \R^n$ is a weight vector (to be computed). Thus the solution is searched in the subspace spanned by the column vectors of $\Phi$. The linear representation helps in two ways. Firstly, given any state $s$, its approximate value can be recovered as $\tj(s)=\phi(s) \tr$ (where $\phi(s)$ is the $s^{th}$ row of $\Phi$). Secondly, a greedy policy $\tilde{u}=u_{\tj}$ (see \Cref{notations}-\eqref{greedy}) can be found using the approximate value function. Note that all computations are $O(k)$ which is independent of the number of states.\par
It is natural to expect that approximations lead to errors and it is important to quantify the errors. For a given ADP method, theoretical performance analysis involves  bounding the error terms $||J^*-\tilde{J}||$  and $\norm{J^*-J_{\tu}}$ which denote the error in approximating the value function, and performance loss due to following policy $\tu$ respectively (where $\norm{\cdot}$ is an appropriate norm). Further, in most cases the error terms reveal some structure that can offer insights and act as guide to the designer of the ADP method (for example the choice of $\Phi$). The focus of this paper is to present the error analysis for generalized reduced linear programming (GRLP), a new ADP formulation which we introduce in \Cref{sec:grlp}.
